Communicative pressures at the semantics-pragmatics interface:\\ Learning biases may prevent the lexicalization of pragmatic inferences

### 2016-09-27 paper/paper.tex
draft analysis section for SIs

\begin{abstract} Certain lexical meanings enable for pragmatic enrichments in a notably productive fashion. This raises the challenge to justify their regular selection, in particular, over alternatives that codify semantically what is conveyed pragmatically. To address this challenge, we propose a general model that integrates iterated Bayesian learning in the replicator-mutator dynamics. This model allows for population-level analyses of the effects of linguistic pressures on probabilistic language users with varied degrees of pragmatic sophistication and distinct languages. We showcase the model's use and predictions in a case study on the (lack of) lexicalization of scalar implicatures. The results suggest simpler semantic representations to be selected for when languages are pressured towards learnability and compression, provided that pragmatic reasoning can compensate for the disadvantage in expressivity that users of such languages otherwise incur.

\section{The semantics-pragmatics divide}\label{sec:introduction}

\end{abstract}

In linguistic theorizing, it is common to draw a distinction between semantics and pragmatics. Broadly speaking, the former concerns the truth-conditional content of expressions, whereas the latter concerns information beyond literal meanings and their composition. An important consequence of this distinction is that the information conveyed by an utterance is seldom, if ever, solely determined by semantics, but rather in tandem with pragmatics.

Much research at the semantics-pragmatics interface has been aimed at characterizing expressions in terms of either domain, or their interplay. However, an issue that has received little attention is the justification of semantic structure in light of pragmatics. The present investigation seeks to fill this gap by analyzing the effects linguistic pressures have on the selection and pervasiveness of particular lexical meanings under consideration of pragmatic enrichments.

In recent years, similar questions about the emergence and change of linguistic features have lead to a surge in models to address them (see \citealt{steels:2015} and \citealt{tamariz+kirby:2016} for recent overviews). Our starting point is given by the overarching argument that has crystalized from accumulated mathematical, experimental and cross-linguistic evidence in this literature: Natural languages need to be well-adapted to communicative needs within a linguistic community, but also need to be learnable to survive their faithful transmission across generations. More succinctly; natural languages are pressured for expressivity as well as learnability.

We build on these insights by modeling these pressures using the replicator-mutator dynamics (see \citealt{hofbauer+sigmund:2003} for an overview). This allows for the inspection of their interaction by combining functional pressure on successful communication with effects of learning biases on (iterated) Bayesian learning \citep{griffiths+kalish:2007}. The semantics-pragmatics distinction and its effect on production and comprehension is made precise by considering probabilistic models of rational language use in populations with distinct lexica \citep{frank+goodman:2012,franke+jaeger:2014, bergen+etal:2016}.

%In this way, the model establishes a link between models of synchronic probabilistic rational language use and diachronic models of cultural evolution.

@@

\section{Simplicity, expressivity, and learnability}

The emergence and change of linguistic structure is influenced by many intertwined factors, ranging from biological and socio-ecological to cultural \citep{steels:2011,tamariz+kirby:2016}. Social and ecological pressures determine communicative needs, while biology determines the architecture that enables and constrains their means of fulfillment. In the following, our focus lies on the latter, cultural factor, wherein processes of linguistic change are understood as shaped by its use and transmission. That is, as a result of cultural evolution.

@@

As noted above, expressivity and learnability are two major competing pressures. Their opposition becomes particularly clear when considering their consequences in the extreme (cf. \citealt{kemp+regier:2012,kirby+etal:2015}). On the one side, a language with a single form is easy to learn but lacking in expressivity for most purposes. On the other, a language that associates a distinct form with all possible meanings its users may want to convey is maximally expressive but challenging to acquire. The most prominent problem that arises from this tension is that of acquiring a language to express a potentially infinite set of meanings through finite means \citep{kirby:2002}. However, this so-called transmission bottleneck is not the only challenge learners confront.

More important for our purposes is the problem of selecting particular hypotheses out of a potentially infinite space of alternatives compatible with the data learners are exposed to. At the semantics-pragmatics interface this concerns the selection between functionally similar, if not identical, lexical meanings.  In the following, we argue an integral part of the answer to be that learners are a priori biased towards simpler, more compressed, lexical representations. This corresponds to the argument that rational learners should prefer simpler over more complex explanations of data \citep{feldman:2000, chater+vitanyi:2003, piantadosi+etal:2012a, kirby+etal:2015,piantadosi+etal:underreview}. In linguistics, a drive for simplicity has been argued to underpin speaker preferences for brevity and ease of articulation, as well as to pressure languages towards lexical ambiguity and grammatical compression \citep{zipf:1949,grice:1975,piantadosi+etal:2012, kirby+etal:2015}. As a broader cognitive principle, the use of simplicity as means to select between hypotheses has a long standing tradition. Crucially, \citet{chater+vitanyi:2003} give a number of compelling arguments for simplicity on both mathematical and empirical grounds.

The remainder of this section introduces the individual components of the model in more detail, as well as the assumptions underlying them. These are: (i) languages and their use, (ii) pressures towards expressivity and learnability, regulated by the replicator and mutator dynamics, respectively, as well as (iii) a bias towards simpler semantic representations, codified as a language learner's prior. After laying out the model, we discuss its application to the lack of lexicalization of scalar implicatures.

\subsection{Languages and linguistic behavior}
Lexica codify the truth-conditions of a language's expressions, i.e., its semantics. A convenient way to represent such lexica is by $(|S|,|M|)$-Boolean matrices, where $S$ is a set of states of affairs, or meanings, to convey and $M$ a set of messages of the language \citep{franke+jaeger:2014}. For instance, the following two lexica fragments determine the truth-conditions of two messages, $m_1$ and $m_2$, for two states, $s_1$ and $s_2$:

@@

In words, according to lexicon $L_a$, $m_1$ is true of state $s_1$ as well as of $s_2$. In contrast, message $m_1$ is only true of $s_1$ in $L_b$. Otherwise, the two languages are truth-conditionally equivalent.

To make the distinction betweeen semantics and pragmatics precise, we distinguish between two kinds of linguistic behavior. {\em Literal interlocutors} produce and interpret messages literally. That is, their linguistic choices are guided by their lexica only. In contrast, {\em pragmatic interlocutors} engage in mutual reasoning to inform their choices. For instance, a rational speaker of $L_a$ who reasons about her addressee should use $m_1$ to signal state $s_1$ given that $s_2$ can unambiguously be conveyed with $m_2$. Analogously, should rational hearers expect their interlocutors to reason along these lines, they will interpret ambiguous $m_1$ accordingly. Note in particular that according to this strenghtening of $m_1$, $L_a$ is indistinguishable from $L_b$ in terms of expressivity if its users are pragmatic reasoners.

Following models of rational language use such as Rational Speech Act models \citep{frank+goodman:2012} and their game-theoretic counterparts \citep{benz+etal:2005a,franke:2009,franke+jaeger:2014}, this kind of signaling behavior is captured by a hierarchy over reasoning types. The hierarchy's bottom, level $0$, corresponds to literal language use. Pragmatic language users of level $n + 1$ behave rationally according to (expected) level $n$ behavior of their interlocutors. The behavior of literal and pragmatic hearers of a language $L$ is given by their respective selection functions in (\ref{h:level0}) and (\ref{h:leveln}). Mutatis mutandis for the speaker functions in (\ref{s:level0}) and (\ref{s:leveln}).

@@

According to (\ref{h:level0}), a literal hearer's interpretation of a message $m$ as a state $s$ depends on her lexicon and her prior over states, $pr \in \Delta(S)$. The literal speaker's choice in (\ref{s:level0}) is regulated by a soft-max parameter $\lambda$, $\lambda \geq 1$ \citep{luce:1959,sutton+barto:1998}. As $\lambda$ increases, choices made in production are more rational in that higher values lead to more deterministic in line with expected utility maximization.

For the most part, pragmatic behavior mirrors its literal counterpart. As described above, their difference lies in that level $n+1$ speakers/hearers reason about level $n$ hearer/speaker behavior instead of solely relaying on their lexicon. That is, they reason about how a rational level $n$ interlocutor would use or interpret a message and behave according to these expectations. Additionally, pragmatic production is further regulated by a parameter $\alpha$ which controls the tension between semantics and pragmatics, $\alpha \in (0,1]$. Lower values lead to more literal production, whereas higher values lead to stronger pragmatic behavior.

The combination of a lexicon with its use, i.e., a particular level of linguistic sophistication, yields a type $t$. Types are the basic units on which our population dynamics operate.

@@

In a nutshell, this explanation posits that scalar implicatures fail to lexicalize because, all else being equal, speakers prefer to communicate as economically as possible and pragmatic reasoning enables them to do so. Compare this with a hypothetical language lexicalizes two expressions instead of a single scalar one; one with and one lacking an upper-bound.\footnote{The observation that monomorphemic expression that lexically rule out stronger alternatives are unattested across languages has received substantial argumentative support (most prominently in \citealt[252-267]{horn:1984} but also e.g. in \citealt{horn:1972,traugott:2004,vdAuwera:2010}). To the best of our knowledge, this claim stands unchallenged.} Should the explanation rest on purely functional grounds, we see four conditions that may pressure languages for English-like semantics over this alternative. First, contextual cues are strongly reliable. Second, morphosyntactic disambiguation is seldom necessary. Third, morphosyntactic disambiguation is only marginally dispreferred. Fourth, larger lexica are costly.

@@

\hl{More  discussion about parameter interaction}

@@

\subsection{Discussion}
changes in sequence length influence the population in a predictable way: smaller values lead to more heterogeneous populations whereas larger ones lead to more pronounced differences. This is expected insofar as the likelihood that a sequence of length $1$ was produced by any type is relatively uniform (modulo prior) whereas the likelihood of types with lexica $L_1$ - $L_3$ to produce, for instance, a sequence of $10$ observations consistently with the same state-message combination is less likely than for pragmatic players using $L_4$ - $L_6$ or literal $L_4$. Thus, while noteworthy, sequence length has no direct bearing on the main contrast of interest. Similar considerations hold for $\alpha$ and $\lambda$ -- set to $1$ and $50$ in the following. Overall, lower rationality in $\lambda$ or more pragmatic violations in $\alpha$ lead to a higher selection of lexica with semantic upper-bounds. The fitness of pragmatic behavior increases with higher $\lambda$-/$\alpha$-values. In other words, these parameters level the functional contrast between $L_4$ and $L_5$.

- There is experimental evidence that scalar implicature calculation costs effort and takes additional processing time (Schaeken, Snedeker, ...).

\section{General discussion}

@@

{\bf Extensions:}

\subparagraph{(I) Cost for pragmatic reasoning.} At least in the CogSci setup the effect of adding cost to pragmatic reasoning is unsurprising: High cost for pragmatic signaling lowers the prevalence of pragmatic types. Lexica that semantically encode an upper-bound benefit the most from this. However, the cost needed to be substantial to make the pragmatic English-like lexicon stop being the incumbent type (particularly when learning is communal).

\subparagraph{(II) Negative learning bias.} Instead of penalizing complex semantics (semantic upper-bounds) one may consider penalizing simple semantics (no upper-bounds). This is useful as a sanity check but also yields unsurprising results in the CogSci setup: The more learners are biased against simple semantics, the more prevalent are lexica that semantically encode upper-bounds.

\subparagraph{(III) Inductive bias.} A second learning bias that codifies the idea that lexica should be uniform, i.e. be biased towards either lexicalizing an upper-bound for all weaker alternatives in a scalar pair or for none.

\subparagraph{(IV)  Uncertainty.} The other advantage of non-upper bounded semantics lies in being non-committal to the negation of stronger alternatives when the speaker is uncertain. Adding this to the model requires the most changes to our present setup and some additional assumptions about the cues available to players to discern the speaker's knowledge about the state she is in.

\subparagraph{(V) More scalar pairs.} Taking into consideration more than one scalar pair. Preliminary results suggest that this does not influence the results in any meaningful way without further additions, e.g. by (III).

\subparagraph{(VI) More lexica.} Not necessary. Preliminary results suggest that considering more lexica has no noteworthy effect on the dynamics (tested with all possible 2x2 lexica).

\subparagraph{(VII) State frequencies.} Variations on state frequencies. This may have an interesting interaction with (III).

\subparagraph{(VIII) Reintroduction of communal learning.} One possibility: The probably $N_{ij}$ with which a child of $t_i$ adopts $t_j$ could be the weighted sum of $Q_{ij}$ (as before) and a vector we get from learning from all of the population: $L_j = \sum_d P(d | \vec{p})  P(t_j | d)$, where $P(d | \vec{p}) = \sum_{i} P(d | t_i)  \vec{p}_i$ is the probability of observing $d$ when learning from a random member of the present population distribution.

@@

>>>>

\begin{abstract} Certain classes of lexical meanings enable for pragmatic enrichments in a notably productive fashion. This raises the challenge to justify their regular selection over alternatives that codify semantically what is conveyed pragmatically. To address this issue, we propose a general model that integrates iterated Bayesian learning in the replicator-mutator dynamics. This model generates predictions about the effects of linguistic pressures on selection and transmission in populations of probabilistic language users with varied degrees of pragmatic sophistication. We apply this model in a case study on the (lack of) lexicalization of scalar implicatures. The results show that simpler semantic representations are selected for when languages are pressured towards learnability and compression, provided that pragmatic reasoning can compensate for the disadvantage in expressivity that users of such languages otherwise incur. We argue this result to shed light on the lack of lexicalization of scalar inferences, as well as the semantics-pragmatics distinction more generally.
\end{abstract}

\section{Simplicity, expressivity, and learnability at the semantics-pragmatics interface}\label{sec:introduction}

In linguistic theorizing it is common to draw a distinction between semantics and pragmatics. Broadly speaking, the former concerns the truth-conditional content of expressions whereas the latter concerns information beyond literal meanings and their composition. An important consequence of this distinction is that the information conveyed by an utterance is seldom, if ever, solely determined by semantics, but rather in tandem with pragmatics. Much research at the semantics-pragmatics interface has been aimed at characterizing expressions in terms of either domain or their interplay. However, an issue that has received little attention is the justification of semantic structure in light of pragmatics. The present investigation seeks to fill this gap by analyzing the effects linguistic pressures have on the selection and pervasiveness of particular lexical meanings under consideration of their possible pragmatic enrichments.

In recent years questions about the emergence and stability of linguistic features have lead to a surge in models that analyze their transmission within and across generations (see \citealt{steels:2015} and \citealt{tamariz+kirby:2016} for recent overviews). Our starting point is given by the overarching argument that has crystalized from the accumulated mathematical, experimental and cross-linguistic evidence in this literature: Natural languages need to be well-adapted to communicative needs within a linguistic community, but also need to be learnable to survive their faithful transmission across generations. More succinctly; natural languages are pressured for expressivity and learnability.

We build on these insights by modeling these pressures using the replicator-mutator dynamics (see \citealt{hofbauer+sigmund:2003} for an overview). This allows for the inspection of their interaction by combining functional pressure on successful communication with effects of learning biases on (iterated) Bayesian learning \citep{griffiths+kalish:2007}. The semantics-pragmatics distinction and its effect on production and comprehension are made precise by considering probabilistic models of rational language use in populations with distinct lexica \citep{frank+goodman:2012,franke+jaeger:2014, bergen+etal:2016}. \hl{SOMETHING MISSING HERE} \hl{In this way, the model establishes a link between models of synchronic probabilistic rational language use and diachronic models of cultural evolution. }

@@

The emergence and change of linguistic structure is influenced by many intertwined factors. These range from biological and socio-ecological to cultural \citep{steels:2011,tamariz+kirby:2016}. Social and ecological pressures determine communicative needs, while biology determines the architecture that enables and constrains the means by which they can be fulfilled. Our focus lies on the latter, cultural, factor wherein processes of linguistic change are understood as shaped by language use and its transmission. That is, as a result of a process of cultural evolution.

@@

As noted above, expressivity and learnability have been identified as two central competing pressures. Their opposition becomes particularly clear when considering their consequences in the extreme (cf. \citealt{kemp+regier:2012,kirby+etal:2015}). On one extreme, a language with a single form is easy to learn but lacking in expressivity. On the other, a language that associates a distinct form with all possible meanings its users may want to convey is maximally expressive but challenging to acquire. The most prominent problem that arises from this tension is that of acquiring a language to express a potentially infinite set of meanings through finite means \citep{kirby:2002}. However, this so-called transmission bottleneck is not the only challenge learners confront. Instead, a problem that is more important for our present purpose is that of selecting particular hypotheses out of a (potentially infinite) space of alternatives compatible with the data learners are exposed to. At the semantics-pragmatics interface this concerns the selection between functionally similar, if not identical, lexical meanings.  In the following, we argue an integral part of the answer to be that learners are a priori biased towards simpler, more compressed, lexical representations. That is, rational learners should prefer simpler over more complex explanations of the data they witness \citep{feldman:2000, chater+vitanyi:2003, piantadosi+etal:2012a, kirby+etal:2015,piantadosi+etal:underreview}. In linguistics, a drive for simplicity has been argued to underpin speaker preferences for brevity and ease of articulation, as well as to pressure languages towards lexical ambiguity and grammatical compression \citep{zipf:1949,grice:1975,piantadosi+etal:2012, kirby+etal:2015}. \hl{As a broader cognitive principle, the use of simplicity as means to select between hypotheses has a long standing tradition. Crucially, \citet{chater+vitanyi:2003} give a number of compelling arguments for simplicity on both mathematical and empirical grounds.}

The remainder of this section introduces the individual components of the model in more detail, as well as the assumptions underlying them. These are: (i) the representation of languages and their use, (ii) pressures towards expressivity and learnability, regulated by the replicator and mutator dynamics, respectively, as well as (iii) a bias towards simpler semantic representations, codified as a language learner's prior. After laying out the model, we discuss its application to the lack of lexicalization of scalar implicatures.

\subsection{Lexica and linguistic behavior}
Lexica codify the truth-conditions of a language's expressions, i.e., its semantics. A convenient way to represent them is by $(|S|,|M|)$-Boolean matrices, where $S$ is a set of states of affairs -- or meanings -- to convey and $M$ a set of messages of the language \citep{franke+jaeger:2014}. For example, the following two fragments determine the truth-conditions of two messages, $m_1$ and $m_2$, for two states, $s_1$ and $s_2$.

@@

According to lexicon $L_a$ the message $m_1$ is true of state $s_1$ as well as of $s_2$. In contrast, $m_1$ is only true of $s_1$ in $L_b$. Otherwise, the two languages are truth-conditionally equivalent.

To make the distinction betweeen semantics and pragmatics precise, we distinguish between two kinds of linguistic behavior. {\em Literal interlocutors} produce and interpret messages literally. That is, their linguistic choices are guided only by their lexica. In contrast, {\em pragmatic interlocutors} engage in mutual reasoning to inform their choices. For instance, a rational speaker of $L_a$ who reasons about her addressee may use $m_1$ to exclusively signal state $s_1$ given that $s_2$ can unambiguously be conveyed through $m_2$. Accordingly, rational hearers that expect their interlocutor to reason along these lines will interpret ambiguous $m_1$ as $s_1$. An important consequence of this pragmatic process is that it renders $L_a$ indistinguishable from $L_b$ in terms of expressivity.

Following models of rational language use such as Rational Speech Act models \citep{frank+goodman:2012} and their game-theoretic counterparts \citep{benz+etal:2005a,franke:2009,franke+jaeger:2014}, this kind of signaling behavior is captured by a resaoning hierarchy. The hierarchy's bottom, level $0$, corresponds to literal language use. Pragmatic language users of level $n + 1$ behave rationally according to (expected) level $n$ behavior of their interlocutors. (\ref{h:level0}) and (\ref{h:leveln}) specify the behavior of literal and pragmatic hearers of a language $L$. Mutatis mutandis for the literal and pragmatic speakers in (\ref{s:level0}) and (\ref{s:leveln}).

@@

According to (\ref{h:level0}) a literal hearer's interpretation of a message $m$ as a state $s$ depends on her lexicon and her prior over states, $pr \in \Delta(S)$. The literal speaker's behavior as given in (\ref{s:level0}) is regulated by a soft-max parameter $\lambda$, $\lambda \geq 1$ \citep{luce:1959,sutton+barto:1998}. As $\lambda$ increases, choices made in production are more rational. That is, higher values lead to behavior that is increasingly in line with expected utility maximization.

Pragmatic behavior is similar to its literal counterpart. As in our informal sketch, their difference lies in that level $n+1$ speakers/hearers reason about level $n$ hearer/speaker behavior instead of solely relaying on their lexicon. That is, they reason about how a rational level $n$ interlocutor would use or interpret a message, and behave according to these expectations. Pragmatic production is further regulated by a parameter $\alpha$ which controls the tension between semantics and pragmatics, $\alpha \in (0,1]$. Lower values lead to more literal production whereas higher values lead to stronger pragmatic behavior.

The combination of a lexicon with its use, i.e., a level in the reasoning hierarchy, is called a type. These are the basic units on which our population dynamics operate.

@@

In a nutshell, this explanation posits that scalar implicatures fail to lexicalize because, all else being equal, speakers prefer to communicate as economically as possible and pragmatic reasoning enables them to do so. Compare this with a hypothetical language lexicalizes two expressions instead of a single scalar one; one with and one lacking an upper-bound. Should the explanation rest on purely functional grounds, we see four conditions that may pressure languages for English-like semantics over this alternative. First, contextual cues are strongly reliable. Second, morphosyntactic disambiguation is seldom necessary. Third, morphosyntactic disambiguation is only marginally dispreferred. Fourth, larger lexica are costly.

@@

\hl{More  discussion about parameter interaction: changes in sequence length influence the population in a predictable way: smaller values lead to more heterogeneous populations whereas larger ones lead to more pronounced differences. This is expected insofar as the likelihood that a sequence of length $1$ was produced by any type is relatively uniform (modulo prior) whereas the likelihood of types with lexica $L_1$ - $L_3$ to produce, for instance, a sequence of $10$ observations consistently with the same state-message combination is less likely than for pragmatic players using $L_4$ - $L_6$ or literal $L_4$. Thus, while noteworthy, sequence length has no direct bearing on the main contrast of interest. Similar considerations hold for $\alpha$ and $\lambda$ -- set to $1$ and $50$ in the following. Overall, lower rationality in $\lambda$ or more pragmatic violations in $\alpha$ lead to a higher selection of lexica with semantic upper-bounds. The fitness of pragmatic behavior increases with higher $\lambda$-/$\alpha$-values. In other words, these parameters level the functional contrast between $L_4$ and $L_5$. }

@@

\subsection{Discussion}
Broadly speaking these results suggest that a lack of semantic upper-bounds coupled with pragmatic reasoning can overcome selective pressures and stabilize in a population provided there is a bias for simpler representations. This outcome is particularly encouraging in light of the other advantages a lack of semantic upper-bounds may confer that were discussed above. 
\hl{The model gives a justification for lexicalization patterns found in natural language, as well as for the failure to lexicalize certain pragmatic inferences. That is, while the puzzle raised by semantics is hard to explain by purely functionalmeans,  we  suggest  part  of  the  answer  to  lie  in  learnability.   Simpler semantic representations are more likely to be learned,  and pragmatic reasoning can counteract functional disadvantages otherwise incurred.  This result is of particular relevance for the longstanding assumption of a divide and interaction between semantics and pragmatics and offers an account of why (certain) pragmatic inferences are not part of the literal meaning of expressions. It furthermore leaves open the possibility of such inferences to fossilize when they do not compete against a lexical simplicity bias}

The model predicts this result to hinge on four conditions. First, types need to be pressured toward both expressivity and learnability. Second, language use can not be too unpredictable; low $\alpha$ or $\lambda$ values render non-upper-bounded lexical meanings more prone to communicative failure and more difficult to learn. Third, learners should prefer simpler over more complex lexical representations. Fourth, the strength by which they need to be preferred depends on the inferential mechanism of learners. That is, to which degree learners maximize the posterior. Put differently, high rationality in learning and choice requires a weaker bias towards simpler representations. The selection of lexical meanings lacking upper-bounds is robust against parameter perturbations under these conditions.

A large range of parameter values lead to the stable incumbency of scalar implicatures. However, while this is predicted by the literature, it is less clear to what proportion other types should be represented. On the one hand, there are  reasons to expect functionally suboptimal types $L_1$-$L_3$ to be largely ruled out. They fail to enable their users to communicate without substantial error both amongst themselves and with other types. In short, these lexica stand at odds with past theoretical and empirical observations. On the other, this is not so for $L_4$.\footnote{$L_6$ presents a special case. In our current setup, it mirrors $L_5$ in allowing for pragmatic strenghtening of message $m_2$ rather than $m_1$. However, its association of $s_1$ with $m_1$ and $s_2$ with $m_2$ under favourable parameter conditions is achieved by ruling out the ``some but not all''-state $s_1$ and not, as with scalar implicatures, the ``all''-state $s_2$. $L_6$ speakers thusly have a ``some''-message strengthened to convey ``all but not some but not all''. The observation that monomorphemic expression that lexically rule out stronger alternatives are unattested across languages has received substantial argumentative support (most prominently in \citealt[252-267]{horn:1984} but also e.g. in \citealt{horn:1972,traugott:2004,vdAuwera:2010}). However, our present setup is blind to these differences.} Notwithstanding, it is possible that natural language communities are composed of such mixed populations or that a single speaker may entertain $L_4$-like semantics for one scalar expression and $L_5$-like semantics for another \hl{CITATIONS FOR THIS PURPOSE}.

Lastly, there are a number of extensions one may want to consider. For instance, the consideration of larger lexica that consider epistemic uncertainty. Another interesting additon relates to possible disadvantages of pragmatic reasoning and the effect of state frequencies on fossilization. We tacitly assumed pragmatic reasoning to come at no cost. However, there is experimental evidence that suggests  that the pragmatic derivation of upper-bounds costs effort and takes additional processing time (cf. \citealt{deNeys+schaeken:2007, huang+snedeker:2009}. This raises the question at which point such usage-based cost undercuts the learnability advantage of simpler semantic representations. As noted above, this might also depend on a given scalar expression and its frequency. Frequently drawn scalar implicatures might fossilize to avoid cost, while infrequent ones could still be computed on-line.

\section{General discussion}
The model gives a justification for lexicalization patterns found in natural language, as well as for the failure to lexicalize certain pragmatic inferences. That is, while the puzzle raised by semantics is hard to explain by purely functional means, we suggest part of the answer to lie in learnability. Simpler semantic representations are more likely to be learned, and pragmatic reasoning can counteract functional disadvantages  otherwise incurred. This result is of particular relevance for the longstanding assumption of a divide and interaction between semantics and pragmatics and offers an account of why (certain) pragmatic inferences are not part of the literal meaning of expressions. It furthermore leaves open the possibility of such inferences to fossilize when they do not compete against a lexical simplicity bias.

A virtue of this model is that it allows for analysis specific modifications and extensions. Straightforward extensions include larger hypothesis spaces, as well as larger or different lexicon fragments. Another worthwhile modification relates to possible disadvantages of pragmatic reasoning. We tacitly assumed pragmatic reasoning to come at no cost. However, there is experimental evidence for the assumption that the pragmatic derivation of upper-bounds costs effort and takes additional processing time (cf. \citealt{deNeys+schaeken:2007, huang+snedeker:2009}). This raises the question at which point such usage-based cost undercuts the learnability advantage of simpler semantic representations.\footnote{In the present setup this modification has a straightforward effect: A penalty for pragmatic signaling lowers the fitness of pragmatic types, to the advantage of literal types. However, the penalty needs to be substantial to counteract the functional advantage pragmatic $L_5$ has over all but $L_4$ together with its learning advantage.}

@@

@@

\hl{The development of natural languages is driven by complex intertwined pressures. Drawing from past insights we put forward a model that allows for a general and malleable integration of core aspects involved in this process.  Chiefly, the model combines functional pressure, iterated Bayesian learning, and probabilistic speaker and hearer models.  In particular,  this analysis addresses longstanding issues concerning
the  semantics-pragmatics  divide.   It  shows  that  when  pressured for learnability and expressivity, the former force drives for simpler semantic representations inasmuch as pragmatics can compensate for them in language use.  Consequently, semantic  patterns  can  be  explained  in  virtue  of  the  linguistic behavior  of  their  users  and  their  complexity.   Furthermore, the ease of acquisition of simpler semantics offers an answer to why natural languages do not lexicalize certain pragmatic inferences.}
